---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```
```{r}
str(av)
```


Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 
```{r}
library(tidyverse)

deaths <- av %>%
  select(URL, Name.Alias, Death1:Death5, Return1:Return5) %>%
  pivot_longer(
    cols = c(Death1:Death5, Return1:Return5),
    names_to = c(".value", "Time"),
    names_pattern = "(Death|Return)(\\d)"
  ) %>%
  mutate(Time = as.numeric(Time)) %>%
  filter(Death != "")
# pivoted the death and return columns for later, late edit

deaths
```
- Dominic's note: A ton of double entries, but that should be fine because they don't really conflict and many characters come back or die several times

Similarly, deal with the returns of characters.
```{r}
returns <- av %>%
  select(URL, Name.Alias, Return1:Return5) %>%
  pivot_longer(cols = Return1:Return5, names_to = "Time", values_to = "Return") %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Return != "")
# added a select, late edit

returns
```
- Dominic's note: similar to deaths, there are a ton of characters that will duplicate due to dying and returning more than once

Based on these datasets calculate the average number of deaths an Avenger suffers. 
```{r}
avg_deaths <- deaths %>%
  filter(Death == "YES") %>%
  group_by(Name.Alias) %>%
  summarise(num_deaths = n()) %>%
  summarise(avg_deaths = mean(num_deaths))

avg_deaths
```
- Dominic's notes: averaging 1.4 deaths per avenger is pretty good. Not super human though.

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.
Dominic's - "2-in-3 chance of returning from first death, but only 50% from a second or third"

### Include the code
Dominic's - 
```{r}
death_return_rates <- deaths %>%
  filter(Death == "YES") %>%
  group_by(Time) %>%
  summarise(
    total_deaths = n(),
    total_returns = sum(Return == "YES", na.rm = TRUE),
    return_rate = round(total_returns / total_deaths, 2)
  )


death_return_rates

```


Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Dominic's answer - Yeah he is exactly right that there is a "2-in-3 chance of returning from first death, but only 50% from a second or third". Looking at the total returns / total deaths shows that he was pretty spot on. Even though the 3rd time a character has returned has a sample size of 2, which I feel like is cheap.

Upload your changes to the repository. Discuss and refine answers as a team.
